### 2024-07-15 ms.qmd
Updates

## Introduction
Behavioural flexibility, **the ability to adapt behaviour to changing environments** (**REFS**), is considered crucial for invasion success [@sol2002behavioural]. In general, species with higher behavioural flexibility are more likely to survive and establish in new habitats (**REFS**) as behaving flexibly enhances their capacity to exploit new resources and face new challenges (**REFS**).

adaptive flexibility hypothesis says that flexibility may be relevant in early and late stages of invassion (Wright et al 2010). But at the same time behavioural flexibility  it is related to relative brain size (Sol et al 2002, 2008), that may be more fixed than we thought (controversy with Wright et al??). In contrast, other hyoptheses predict this is constant, or there is individual variation in beh flexibility (Wright et al 2010). Since some aspects of early environment can play a prominent role in shaping neural substrate, we hypothesize that beh flexibility may be affected by early environment. Also, since beh flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations. In other words, there may be some mechanisms in invassive species to avoid early developmental constraints. CORT can be essential in this process as it is associated to vertebrates response to environmental changes, and it is known to impact brain development (Zhu et al 2004). Also, temperature may be relevant, since it affects brain development in ectotherms, it can be one of the main environmental changes to face by animals, andvertebrates reponse to changes in temeprature is also associated with the production of CORT (Crino et al 2023).

Here, we aim to study the effects of prenatal environment on behavioural flexibility in the invassion ecology context. To do so, we employed as our model species the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). These two species are closely related and share similar life history traits [@chapple_know_2011; @chapple_biology_2014], but only _L. delicata_ has been succesful in colonising different areas around the globe (**REFS**). Some studies have shown differences in personality traits between these two species that can be related to higher behavioural flexibility in _L. delicata_ (**REFS**), and are likely key for its invassive success (**REFS**). Our goal in this study was to investigate if behavioural flexibility is affected by the prenatal environment and the species specific response to it. Particularly, we were interested in the interactive effects of prenatal temperature and Corticosterone (CORT) - the main type of GC in reptiles [@Crino_2023] - because CORT is usually associated to vertebrates response to changes in temeprature [@Crino_2023], and both factors are known to impact brain development [@coomber_independent_1997; @zhu_prenatal_2004]. We hypothesized that, since behavioural flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations. In other words, there may be some mechanisms in invassive species to avoid early

@@

#### Experimental design
We employed juveniles of both species, _L. delicata_ and _L. guichenoti_ that came from a breeding colony kept in the lab (details of breeding colony and husbandry in Supplementary Material). To see the effect of prenatal CORT and temperature, eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO ('CORT' treatment) or the same volume of the vehicle ('Control'), and were then incubated at different temperatures ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) until hatching.  
To estimate behavioural flexibility, we examined lizards' performance in a reversal learning task using colour as the main conditioned stimulus and food (frozen small *A. domestica*) as a reward (see @fig-Methods). The learning test was divided in three main phases: there was first a habituation phase where lizards had to learn to eat from white 3D-printed PLA ramps (9 L x 4 W x 5 H cm); then a second phase where lizards had to associate the attainability of food with one colour; and finally the reversal task where the colour associated with the attainable food was changed (@fig-Methods B). In associative and reversal tasks we used  using three different ramps that were identical in size and shape to the ones for the habituation phase. The colours of the ramps were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. To control for potential colour biases, we assigned one group of lizards to associate the blue ramp with the attainable food in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. To avoid the use of chemical cues, we placed the food reward in all three ramps, but in two of them the cricket was covered by small white lids with holes on the top. Lizards were tested once daily between 11 to 12 am, when lizards are active. The training phase lasted 38 days, the Associaive learning task 35, and the Reversal learning task 40 days. We recorded and analysed manually all trials form the learning phases using a standard video player (IINA). PR was the main researcher analysing the data, and was blinded to the lizards' treatment until the end of the analyses. We used as the main variable whether the animal chose the correct ramp first or not. We considered that a choice was made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses. Here, we analysed the data of the reversal as a measure of behavioural flexibility. As such, we only included in our analyses those individuals who passed a learning criterion of 80% correct choices in the last 10 trials of the associative task.

```{r, fig-Methods}
#| label: fig-Methods
#| fig.cap: "Experimental design of early environment manipulation (**A**) and learning tasks (**B**). Stages 1-3 indicate the different phases of the habituation process. In the associative and reversal tasks, white lids show the ramps where the food reward was not attainable."

knitr::include_graphics("./Others/BEHFLEX_FIG_1.svg")

>>>>

## Introduction

@@

#### Husbandry  
*Breeding colony* -- The study was conducted with juveniles of the Delicate skink (_Lamprophpolis delicata_) from a breeding colony kept in the lab since 2019. The colony is composed by a total of 270 adults housed in big containers (41.5 L x 30.5 W x 21 H cm) with six lizards (2 males and 4 females) per enclosure. Enclosures are provided with non-stick matting, shelter, and several small water dishes. Water is given daily, and they are fed approx. 40 mid-size crickets (*Acheta domestica*) per enclosure three days a week. Crickets are dusted with calcium weekly and multivitamin and calcium biweekly. To ensure a temperature gradient, we employ a heat chord and a heat lamp following a 12 h light:12 h dark cycle. Room temperatures are set to 22-24 Celsius, and warm side of enclosures is usually at 32 Celsius.

*Eggs collection and incubation* -- Between mid-October 2022 to the end of February 2023, we provided females with a place to lay the eggs by means of small boxes (12.5 L x 8.3 W x 5 H cm) with moist vermiculite inside, that were placed in one extreme of the communal enclosures (see above). We checked for the presence of eggs in the boxes three days a week. After collection, we measured length and width of eggs with a digital caliper to the nearest 0.1 mm and weight them with a (OHAUS, Model spx123) digital scale ± 0.001g error. Then eggs were treated with CORT or vehicle (see CORT and Temperature manipulation below) and were placed in individual cups (80 mL) with moist vermiculite (12 parts water to 4 parts vermiculite). The cups were covered with cling wrap to retain moisture and left in LATWIT 2X5D-R1160 incubators at two different temperatures (see CORT and Temperature manipulation below) until hatching.

*Hatchlings* -- Eggs in the incubator were checked three times a week for hatchlings. After hatchling, we measured juveniles' SVL and Tail Length (TL) with a rule to the nearest mm and weighted them with a (OHAUS, Model spx123) digital scale ± 0.001g error. We then placed hatchlings in individual enclosures (18.7L x 13.2W x 6.3H cm) and provided them with non-stick matting and a small water dish. During this period, they were sprayed water every day and received 3-6 small *A. domestica* crickets three times a week. All care otherwise follows similar protocols to adults (see above).  
Two weeks before we started the training phase (see below), lizards were moved to the experimental arena for acclimatation. The arenas were individual medium size (41 L x 29.7 W x 22 H cm) plastic containers with a shelter (9 L x 6 W x 1.5 H cm) on one of the extremes and a water dish on the other. These new enclosures were placed in two rooms in 7 different racks associated to 7 different CCTV systems (device model DVR-HP210475) that allowed us to record their behaviour during the experiment (see details below). The number of lizards per species and treatment in each rack was counterbalanced to control for any effect of the room or the position of the lizard in the rack. During acclimatation and all the experiment, lizards were fed with only one cricket per day dusted with calcium and multivitamin (see protocol below), and water was supplied *ad libitum*. We provided a temperature gradient by means of a heat cord and heat lamps in a 12 h light: 12 h dark cycle. The rooms temperature was set to between 22-24 Celsius.

#### Learning

```{r, fig-Methods}

```{r, fig-Methods}
#| label: fig-Methods
#| fig.cap: "Panel **A** shows the experimental design of the study. On top, the treatments applied to the eggs. On the bottom the learning task and the brain region extracted together with the physiological analyses performed. In panel **B** the details and measurements of the spatial learning maze."

knitr::include_graphics("./Others/SPAL_METH.svg")
